Allocation Removal by Partial Evaluation in a Tracing JIT

### 2011-01-11 talk/pepm2011/presentation/talk.tex
tweaks

\usepackage[english]{babel}
\usepackage{listings}

@@

\begin{frame}
  \frametitle{Dynamic Languages are Slow: Example}
  Evaluate \texttt{x = a + b; y = x + c} in an interpreter:
  \pause
  \begin{enumerate}
      \item What's the type of a? \texttt{Integer}
      \item What's the type of b? \texttt{Integer}
  \pause
      \item unbox a
      \item unbox b
      \item compute the sum
      \item box the result
      \item store into x
  \pause
      \item What's the type of x? \texttt{Integer}
      \item What's the type of c? \texttt{Integer}
  \pause
      \item unbox x
      \item unbox c
      \item compute the sum
      \item box the result
      \item store into y
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{What to do?}
  \begin{itemize}
      \item Hard to improve in an interpreter
      \item Use a JIT compiler
      \item \textbf{Add a optimization that can deal with heap operations}
  \end{itemize}
\end{frame}

@@

@@

@@

\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
      \item straightforward interpreter can be efficient, given enough technology
      \item successful application of partial evaluation
      \item Prolog can benefit from dynamic compilation
  \end{itemize}
  \pause
  \begin{block}{Future}
      \begin{itemize}
          \item Scale up to larger programs
          \item need some optimization on the interpreter level – indexing
          \item investigate memory usage
      \end{itemize}
  \end{block}
\end{frame}

>>>>

\usepackage[english]{babel}
\usepackage{listings}
\usepackage{ulem}
\usepackage{alltt}

@@

\begin{frame}
  \frametitle{Dynamic Languages are Slow: Example}
  \texttt{x = a + b; y = x + c}

  evaluated in an interpreter:
  \pause
  \begin{enumerate}
      \item What's the type of a? \texttt{Integer}
      \item What's the type of b? \texttt{Integer}
  \pause
      \item unbox a
      \item unbox b
      \item compute the sum
      \item box the result as an \texttt{Integer}
      \item store into x
  \pause
      \item What's the type of x? \texttt{Integer}
      \item What's the type of c? \texttt{Integer}
  \pause
      \item unbox x
      \item unbox c
      \item compute the sum
      \item box the result as an \texttt{Integer}
      \item store into y
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{What to do?}
  \begin{itemize}
      \item Hard to improve in an interpreter
      \item Use a JIT compiler
      \item \textbf{Add a optimization that can deal with heap operations}
      \pause
      \item optimize short-lived objects
  \end{itemize}
\end{frame}

@@

\begin{frame}[containsverbatim]
  \frametitle{Example Trace}
  Trace of \texttt{x = a + b; y = x + c}:
\begin{verbatim}
guard_class(a, Integer)
guard_class(b, Integer)
i1 = get(a, intval)
i2 = get(b, intval)
i3 = int_add(i1, i2)
x = new(Integer)
set(x, intval, i3)
\end{verbatim}
\begin{verbatim}
guard_class(x, Integer)
guard_class(c, Integer)
i4 = get(x, intval)
i5 = get(c, intval)
i6 = int_add(i4, i5)
y = new(Integer)
set(y, intval, i6)
\end{verbatim}
\end{frame}

@@

\begin{frame}[containsverbatim]
  \frametitle{Optimized Example Trace}
  Trace of \texttt{x = a + b; y = x + c}:
\begin{alltt}
guard_class(a, Integer)
guard_class(b, Integer)
i1 = get(a, intval)
i2 = get(b, intval)
i3 = int_add(i1, i2)
\sout{x = new(Integer)}
\sout{set(x, intval, i3)}
\sout{guard_class(x, Integer)}
guard_class(c, Integer)
\sout{i4 = get(x, intval)}
i5 = get(c, intval)
i6 = int_add(i4, \emph{i3})
y = new(Integer)
set(y, intval, i6)
\end{alltt}
\end{frame}

@@

\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
      \item very simple partial-evaluation-based
      \item successful application of partial evaluation
      \item Prolog can benefit from dynamic compilation
  \end{itemize}
  \pause
  \begin{block}{Future}
      \begin{itemize}
          \item Scale up to larger programs
          \item need some optimization on the interpreter level – indexing
          \item investigate memory usage
      \end{itemize}
  \end{block}
\end{frame}
